---
title: "Hands-on_Ex04"
author: "Hin Phung"
editor: visual
---

# 9.0 Visual Statistical Analysis

### 9.1 Learning Outcome

The following are the learning outcome:

-   ggstatsplot package to create visual graphics with rich statistical information,

-   performance package to visualise model diagnostics, and

-   parameters package to visualise model parameters

### 9.2 Visual Statistical Anaysis with ggstatplot

ggstatsplot is an extension of ggplot2 package for creating graphics with details from statistical tests included in the information-rich plots themselves.It provides alternative statistical inference methods by default. To follow best practices for statistical reporting.

### 9.3 Getting Started

#### 9.3.1 Installing and launching R packages

```{r}
pacman::p_load(ggstatsplot, tidyverse)

```

#### 9.3.2 Importing Data

```{r}
#| warning: false
exam_data <- read_csv("data/Exam_data.csv")

```

#### 9.3.3 One-sample test: gghistostats() method

In the code chunk below, gghistostats() is used to to build an visual of one-sample test on English scores.

```{r}
set.seed(1234)

gghistostats(
  data = exam_data,
  x = ENGLISH,
  type = "bayes",
  test.value = 60,
  xlab = "English scores"
)

```

Default information: - statistical details - Bayes Factor - sample sizes - distribution summary

#### 9.3.4 Unpacking the Bayes Factor

A Bayes factor is the ratio of the likelihood of one particular hypothesis to the likelihood of another. It can be interpreted as a measure of the strength of evidence in favor of one theory among two competing theories.

That's because the Bayes factor gives us a way to evaluate the data in favor of a null hypothesis, and to use external information to do so. It tells us what the weight of the evidence is in favor of a given hypothesis.

When we are comparing two hypotheses, H1 (the alternate hypothesis) and H0 (the null hypothesis), the Bayes Factor is often written as B10.

The Schwarz criterion is one of the easiest ways to calculate rough approximation of the Bayes Factor.

#### 9.3.5 How to interpret Bayes Factor

A Bayes Factor can be any positive number. One of the most common interpretations is this one---first proposed by Harold Jeffereys (1961) and slightly modified by Lee and Wagenmakers in 2013.

#### 9.3.6 Two-sample mean test: ggbetweenstats()

In the code chunk below, ggbetweenstats() is used to build a visual for two-sample mean test of Maths scores by gender.

```{r}
ggbetweenstats(
  data = exam_data,
  x = GENDER, 
  y = MATHS,
  type = "np",
  messages = FALSE
)

```

Default information: - statistical details - Bayes Factor - sample sizes - distribution summary

#### 9.3.7 Oneway ANOVA Test: ggbetweenstats() method

In the code chunk below, ggbetweenstats() is used to build a visual for One-way ANOVA test on English score by race.

```{r}
ggbetweenstats(
  data = exam_data,
  x = RACE, 
  y = ENGLISH,
  type = "p",
  mean.ci = TRUE, 
  pairwise.comparisons = TRUE, 
  pairwise.display = "s",
  p.adjust.method = "fdr",
  messages = FALSE
)
```

\-"ns" → only non-significant

\-"s" → only significant -"all" → everything

##### 9.3.7.1 ggbetweenstats - Summary of tests

#### 9.3.8 Significant Test of Correlation: ggscatterstats()

In the code chunk below, ggscatterstats() is used to build a visual for Significant Test of Correlation between Maths scores and English scores.

```{r}
ggscatterstats(
  data = exam_data,
  x = MATHS,
  y = ENGLISH,
  marginal = FALSE,
  )

```

#### 9.3.9 Significant Test of Association (Dependence): ggbarstats() methods

In the code chunk below, the Maths scores is binned into a 4-class variable by using cut().

```{r}
exam1 <- exam_data %>% 
  mutate(MATHS_bins = 
           cut(MATHS, 
               breaks = c(0,60,75,85,100))
)
```

In this code chunk below ggbarstats() is used to build a visual for Significant Test of Association

```{r}
ggbarstats(exam1, 
           x = MATHS_bins, 
           y = GENDER)
```

### 9.4 Visualising Models

In this section, you will learn how to visualise model diagnostic and model parameters by using parameters package.

Toyota Corolla case study will be used. The purpose of study is to build a model to discover factors affecting prices of used-cars by taking into consideration a set of explanatory variables.

### 9.5 Getting Started

### 9.6 Installing and loading the required libraries

### 

```{r}
pacman::p_load(readxl, performance, parameters, see)
```

#### 9.6.1 Importing Excel file: readxl methods

In the code chunk below, read_xls() of readxl package is used to import the data worksheet of ToyotaCorolla.xls workbook into R.

```{r}
#| warning: False
car_resale <- read_xls("data/ToyotaCorolla.xls", 
                       "data")
car_resale
```

#### 9.6.2 Multiple Regresssion Model using Im()

The code chunk below is used to calibrate a multiple linear regression model by using lm() of Base Stats of R.

```{r}
model <- lm(Price ~ Age_08_04 + Mfg_Year + KM + 
              Weight + Guarantee_Period, data = car_resale)
model
```

#### 9.6.3 Model Diagnostic: checking for multicollinearity

In the code chunk, check_collinearity() of performance package.

```{r}
check_collinearity(model)
```

```{r}
check_c <- check_collinearity(model)
plot(check_c)
```

#### 9.6.4 Model Diagnostic: checking normality assumption

In the code chunk, check_normality() of performance package.

```{r}
model1 <- lm(Price ~ Age_08_04 + KM + 
              Weight + Guarantee_Period, data = car_resale)
```

```{r}
check_n <- check_normality(model1)

```

```{r}
plot(check_n)

```

#### 9.6.5 Model Diagnostic: check model for homogeneity of variances

In the code chunk, check_heteroscedasticity() of performance package.

```{r}
check_h <- check_heteroscedasticity(model1)

```

```{r}
plot(check_h)

```

#### 9.6.6 Model Diagnostic: complete check

We can also perform the complete by using check_model().

```{r}
check_model(model1)

```

#### 9.6.7 Visualising Regression Parameters: see methods

In the code below, plot() of see package and parameters() of parameters package is used to visualise the parameters of a regression model.

```{r}
plot(parameters(model1))

```

#### 9.6.8 Visualising Regression Parameters: ggcoefstats() methods

In the code below, ggcoefstats() of ggstatsplot package to visualise the parameters of a regression model.

```{r}
ggcoefstats(model1, 
            output = "plot")

```

# 10. Visualising Uncertainty

### 10.1 Learning Outcome

### 10.2 Visualizing the Uncertainty of point estimates

-   A point estimate is a single number, such as a mean.

-   Uncertainty is expressed as standard error, confidence interval, or credible interval

-   Important:Don't confuse the uncertainty of a point estimate with the variation in the sample

```{r}
pacman::p_load(tidyverse, plotly, crosstalk, DT, ggdist, gganimate)

```

```{r}
exam <- read_csv("data/Exam_data.csv")

```

#### 10.2.1 Visualizing the uncertainty of point estimates: ggplot2 methods

The code chunk below performs the followings:

-   group the observation by RACE,

-   computes the count of observations, mean, standard deviation and standard error of Maths by RACE, and

-   save the output as a tibble data table called my_sum.

```{r}
my_sum <- exam %>%
  group_by(RACE) %>%
  summarise(
    n=n(),
    mean=mean(MATHS),
    sd=sd(MATHS)
    ) %>%
  mutate(se=sd/sqrt(n-1))

```

```{r}
knitr::kable(head(my_sum), format = 'html')

```

#### 10.2.2 Visualizing the uncertainty of point estimates: ggplot2 methods

The code chunk below is used to reveal the standard error of mean maths score by race.

```{r}
ggplot(my_sum) +
  geom_errorbar(
    aes(x=RACE, 
        ymin=mean-se, 
        ymax=mean+se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    size=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  ggtitle("Standard error of mean 
          maths score by race")
```

#### 10.2.3 Visualizing the uncertainty of point estimates: ggplot2 methods

Plot the 95% confidence interval of mean maths score by race. The error bars should be sorted by the average maths scores.

```{r}
my_sum$RACE = with(my_sum, reorder(RACE,-mean))

ggplot(my_sum) +
  geom_errorbar(
    aes(x=RACE, 
        ymin=mean-1.96*se, 
        ymax=mean+1.96*se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    size=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  ggtitle("95% cobfidence interval of mean 
          maths score by race")

```

#### 10.2.4 Visualising the uncertainty of point estimates with interactive error bars

Plot interactive error bars for the 99% confidence interval of mean maths score by race.


```{r}

My_sum2 <- rename(my_sum, c("No. of pupils" = n, "Ave Scores" = mean, "Std Dev" = sd, "Std Error" = se))

My_sum2

```

```{r}
DT::datatable(My_sum2, class= "compact")

```

`
```{r}


d <- highlight_key(My_sum2) 
p <- ggplot(my_sum) +
  geom_errorbar(
    aes(x=RACE, 
        ymin=mean-2.58*se, 
        ymax=mean+2.58*se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    size=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  ggtitle("95% cobfidence interval of mean 
          maths score by race")

gg <- highlight(ggplotly(p),        
                "plotly_selected")  

crosstalk::bscols(gg,               
                  DT::datatable(d), 
                  widths = 5)        

```




### 10.3 Visualising Uncertainty: ggdist package

-   ggdist is an R package that provides a flexible set of ggplot2 geoms and stats designed especially for visualising distributions and uncertainty.

-   It is designed for both frequentist and Bayesian uncertainty visualization, taking the view that uncertainty visualization can be unified through the perspective of distribution visualization:

    -   for frequentist models, one visualises confidence distributions or bootstrap distributions (see vignette("freq-uncertainty-vis"));
    -   for Bayesian models, one visualises probability distributions (see the tidybayes package, which builds on top of ggdist).

#### 10.3.1 Visualising the certainty of point estimates: ggdist methods

In the code chunk below, stat_pointinterval() of ggdist is used to build a visual for displaying distribution of maths scores by race.

```{r}
exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_pointinterval() +   #<<
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Mean Point + Multiple-interval plot")

```

Gentle advice: This function comes with many arguments, it is advised to read the syntax reference for more detail.

```{r}
exam %>%
  ggplot(aes(x = RACE, y = MATHS)) +
  stat_pointinterval(.width = 0.95,
  .point = median,
  .interval = qi) +
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Mean Point + Multiple-interval plot")

```

#### 10.3.2 Visualising the certainty of point estimates: ggdist methods

Makeover the plot on previous slide by showing 95% and 99% confidence intervals.

```{r}
exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_pointinterval(
    show.legend = FALSE) +   
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Mean Point + Multiple-interval plot")

```






#### 10.3.3 Visualising the certainty of point estimates: ggdist methods

In the code chunk below, stat_gradientinterval() of ggdist is used to build a visual for displaying distribution of maths scores by race.

```{r}
#| warning: false
exam %>%
  ggplot(aes(x = RACE, 
             y = MATHS)) +
  stat_gradientinterval(   
    fill = "skyblue",      
    show.legend = TRUE     
  ) +                        
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Gradient + interval plot")

```

Gentle advice: This function comes with many arguments, it is advised to read the syntax reference for more details.

### 10.4 Visualisaing Uncertainty with Hypothetical Outcome Plots (HOPs)

Step 1: Installing ungeviz package

```{r}
devtools::install_github("wilkelab/ungeviz")

```

Note: Only need to perform this step once.

Step 2: Launch the application in R

```{r}
library(ungeviz)

```

```{r}
#| warning: false
ggplot(data = exam, 
       (aes(x = factor(RACE), y = MATHS))) +
  geom_point(position = position_jitter(
    height = 0.3, width = 0.05), 
    size = 0.4, color = "#0072B2", alpha = 1/2) +
  geom_hpline(data = sampler(25, group = RACE), height = 0.6, color = "#D55E00") +
  theme_bw() + 
  # `.draw` is a generated column indicating the sample draw
  transition_states(.draw, 1, 3)

```

### 10.5 Visualisaing Uncertainty with Hypothetical Outcome Plots (HOPs)

```{r}
#| warning: false
ggplot(data = exam, 
       (aes(x = factor(RACE), 
            y = MATHS))) +
  geom_point(position = position_jitter(
    height = 0.3, 
    width = 0.05), 
    size = 0.4, 
    color = "#0072B2", 
    alpha = 1/2) +
  geom_hpline(data = sampler(25, 
                             group = RACE), 
              height = 0.6, 
              color = "#D55E00") +
  theme_bw() + 
  transition_states(.draw, 1, 3)

```

# 11. Funnel Plots for Fair Comparisons

### 11.1 Overview

Funnel plot is a specially designed data visualisation for conducting unbiased comparison between outlets, stores or business entities. By the end of this hands-on exercise, you will gain hands-on experience on:

-   plotting funnel plots by using funnelPlotR package,

-   plotting static funnel plot by using ggplot2 package, and

-   plotting interactive funnel plot by using both plotly R and ggplot2 packages.

### 11.2 Installing and Launching R Packages

Five R packages will be used. They are:

-   readr for importing csv into R.

-   FunnelPlotR for creating funnel plot.

-   ggplot2 for creating funnel plot manually.

-   knitr for building static html table.

-   plotly for creating interactive funnel plot.

```{r}
pacman::p_load(tidyverse, FunnelPlotR, plotly, knitr)
```

### 11.3 Importing Data

In this section, COVID-19_DKI_Jakarta will be used. The data was downloaded from Open Data Covid-19 Provinsi DKI Jakarta portal. We are going to compare the cumulative COVID-19 cases and death by sub-district (i.e. kelurahan) as at 31st July 2021, DKI Jakarta.

The code chunk below imports the data into R and save it into a tibble data frame object called covid19.

```{r}
covid19 <- read_csv("data/COVID-19_DKI_Jakarta.csv") %>%
  mutate_if(is.character, as.factor)
```

### 11.4 FunnelPlotR methods

FunnelPlotR package uses ggplot to generate funnel plots. It requires a numerator (events of interest), denominator (population to be considered) and group. The key arguments selected for customisation are:

-   limit: plot limits (95 or 99).

-   label_outliers: to label outliers (true or false).

-   Poisson_limits: to add Poisson limits to the plot.

-   OD_adjust: to add overdispersed limits to the plot.

-   xrange and yrange: to specify the range to display for axes, acts like a zoom function.

-   Other aesthetic components such as graph title, axis labels etc.

#### 11.4.1 FunnelPlotR methods: The basic plot

The code chunk below plots a funnel plot.

```{r}
funnel_plot(
  numerator = covid19$Positive,
  denominator = covid19$Death,
  group = covid19$`Sub-district`
)

```

Things to learn from the code chunk above.

-   group in this function is different from the scatterplot. Here, it defines the level of the points to be plotted i.e. Sub-district, District or City. If Cityc is chosen, there are only six data points.

-   By default, data_typeargument is "SR".

-   limit: Plot limits, accepted values are: 95 or 99, corresponding to 95% or 99.8% quantiles of the distribution.

#### 11.4.2 FunnelPlotR methods: Makeover 1

The code chunk below plots a funnel plot.

```{r}
funnel_plot(
  numerator = covid19$Death,
  denominator = covid19$Positive,
  group = covid19$`Sub-district`,
  data_type = "PR",     #<<
  xrange = c(0, 6500),  #<<
  yrange = c(0, 0.05)   #<<
)

```

Things to learn from the code chunk above. + data_type argument is used to change from default "SR" to "PR" (i.e. proportions). + xrange and yrange are used to set the range of x-axis and y-axis

#### 11.4.3 FunnelPlotR methods: Makeover 2

The code chunk below plots a funnel plot.

```{r}
funnel_plot(
  numerator = covid19$Death,
  denominator = covid19$Positive,
  group = covid19$`Sub-district`,
  data_type = "PR",   
  xrange = c(0, 6500),  
  yrange = c(0, 0.05),
  label = NA,
  title = "Cumulative COVID-19 Fatality Rate by Cumulative Total Number of COVID-19 Positive Cases", #<<           
  x_label = "Cumulative COVID-19 Positive Cases", #<<
  y_label = "Cumulative Fatality Rate"  #<<
)

```

Things to learn from the code chunk above.

-   label = NA argument is to removed the default label outliers feature.

-   title argument is used to add plot title.

-   x_label and y_label arguments are used to add/edit x-axis and y-axis titles.

### 11.5 Funnel Plot for Fair Visual Comparison: ggplot2 methods

In this section, we can gain hands-on experience on building funnel plots step-by-step by using ggplot2. It aims to enhance you working experience of ggplot2 to customise speciallised data visualisation like funnel plot.

#### 11.5.1 Computing the basic derived fields

To plot the funnel plot from scratch, we need to derive cumulative death rate and standard error of cumulative death rate.

```{r}
df <- covid19 %>%
  mutate(rate = Death / Positive) %>%
  mutate(rate.se = sqrt((rate*(1-rate)) / (Positive))) %>%
  filter(rate > 0)

```

Next, the fit.mean is computed by using the code chunk below.

```{r}
fit.mean <- weighted.mean(df$rate, 1/df$rate.se^2)

```

#### 11.5.2 Calculate lower and upper limits for 95% and 99.9% CI

```{r}
number.seq <- seq(1, max(df$Positive), 1)
number.ll95 <- fit.mean - 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul95 <- fit.mean + 1.96 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ll999 <- fit.mean - 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
number.ul999 <- fit.mean + 3.29 * sqrt((fit.mean*(1-fit.mean)) / (number.seq)) 
dfCI <- data.frame(number.ll95, number.ul95, number.ll999, 
                   number.ul999, number.seq, fit.mean)

```

#### 11.5.3 Plotting a static funnel plot

In the code chunk below, ggplot2 functions are used to plot a static funnel plot.

```{r}
#| warning: false
p <- ggplot(df, aes(x = Positive, y = rate)) +
  geom_point(aes(label=`Sub-district`), 
             alpha=0.4) +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul95), 
            size = 0.4, 
            colour = "grey40", 
            linetype = "dashed") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ll999), 
            size = 0.4, 
            colour = "grey40") +
  geom_line(data = dfCI, 
            aes(x = number.seq, 
                y = number.ul999), 
            size = 0.4, 
            colour = "grey40") +
  geom_hline(data = dfCI, 
             aes(yintercept = fit.mean), 
             size = 0.4, 
             colour = "grey40") +
  coord_cartesian(ylim=c(0,0.05)) +
  annotate("text", x = 1, y = -0.13, label = "95%", size = 3, colour = "grey40") + 
  annotate("text", x = 4.5, y = -0.18, label = "99%", size = 3, colour = "grey40") + 
  ggtitle("Cumulative Fatality Rate by Cumulative Number of COVID-19 Cases") +
  xlab("Cumulative Number of COVID-19 Cases") + 
  ylab("Cumulative Fatality Rate") +
  theme_light() +
  theme(plot.title = element_text(size=12),
        legend.position = c(0.91,0.85), 
        legend.title = element_text(size=7),
        legend.text = element_text(size=7),
        legend.background = element_rect(colour = "grey60", linetype = "dotted"),
        legend.key.height = unit(0.3, "cm"))
p

```

#### 11.5.4 Interactive Funnel Plot: Plotly + ggplot2

The funnel plot created using ggplot2 functions can be made interactive with ggplotly() of plotly r package.

```{r}
fp_ggplotly <- ggplotly(p,
  tooltip = c("label", 
              "x", 
              "y"))
fp_ggplotly

```




##########################################################################

#### 10.2.3 Visualizing the uncertainty of point estimates: ggplot2 methods

Plot the 95% confidence interval of mean maths score by race. The error bars should be sorted by the average maths scores.

```{r}

my_sum$RACE = with(my_sum, reorder(RACE,-mean))

ggplot(my_sum) +
  geom_errorbar(
    aes(x=RACE, 
        ymin=mean-1.96*se, 
        ymax=mean+1.96*se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    size=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  ggtitle("95% cobfidence interval of mean 
          maths score by race")


```




#### 10.2.4 Visualising the uncertainty of point estimates with interactive error bars

Plot interactive error bars for the 99% confidence interval of mean maths score by race.

```{r}

My_sum2 <- rename(my_sum, c("No. of pupils" = n, "Ave Scores" = mean, "Std Dev" = sd, "Std Error" = se))

My_sum2

```

```{r}
DT::datatable(My_sum2, class= "compact")

```

`
```{r}


d <- highlight_key(My_sum2) 
p <- ggplot(my_sum) +
  geom_errorbar(
    aes(x=RACE, 
        ymin=mean-2.58*se, 
        ymax=mean+2.58*se), 
    width=0.2, 
    colour="black", 
    alpha=0.9, 
    size=0.5) +
  geom_point(aes
           (x=RACE, 
            y=mean), 
           stat="identity", 
           color="red",
           size = 1.5,
           alpha=1) +
  ggtitle("95% cobfidence interval of mean 
          maths score by race")

gg <- highlight(ggplotly(p),        
                "plotly_selected")  

crosstalk::bscols(gg,               
                  DT::datatable(d), 
                  widths = 5)        

```





#### 10.3.2 Visualising the certainty of point estimates: ggdist methods

Makeover the plot on previous slide by showing 95% and 99% confidence intervals.


```{r}
exam %>%
  ggplot(aes(x = RACE, y = MATHS)) +
  stat_pointinterval(.width = 0.95,
                     .point = median,
                     .interval = "confidence",
                     level = 0.95) +
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Mean Point + Multiple-interval plot (95% confidence intervals)"
  )
```


```{r}
exam %>%
  ggplot(aes(x = RACE, y = MATHS)) +
  stat_pointinterval(.width = 0.95,
                     .point = median,
                     .interval = "confidence",
                     level = 0.99) +
  labs(
    title = "Visualising confidence intervals of mean math score",
    subtitle = "Mean Point + Multiple-interval plot (99% confidence intervals)"
  )
```

