---
title: "In-Class Exercise 4"
author: "Hin Phung"
date: "6 May 2023"
execute:
  echo: true
  eval: true
  warning: false
editor: visual
---

```{r}
pacman::p_load(rstatix, gt, patchwork, tidyverse)

```

```{r}
exam_data <-read_csv("data/Exam_data.csv")

```

Plot Q-Q plot. 
Conversely, if the points deviate significantly from straight diagonal line, it is less likely that data is normally distributed.

```{r}
ggplot(exam_data,
       aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()

```


Create a tab to toggle between code chunk and plot

:::panel-tabset

## The plot
```{r}
#| echo: false
#| fig-width: 4
#| fig-height: 4
ggplot(exam_data,
       aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()

```

## The code chunk
```{r}
#| eval: false
#| fig-width: 4
#| fig-height: 4
ggplot(exam_data,
       aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()

```

:::

Create a call-out as notes.

::: {.callout-note}
We can see that the points deviate significant from the straight diagonal line. This is a clear indication thatr the data is not normally distributed.
:::


:::panel-tabset

## The plot
```{r}
#| echo: FALSE
qq <- ggplot(exam_data,
       aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()
  
sw_t <- exam_data %>%
   shapiro_test(ENGLISH) %>%
   gt()
 
tmp <- tempfile(fileext = '.png')
gtsave(sw_t, tmp)
table_png <- png::readPNG(tmp, native= TRUE)
 
qq + table_png
 

```

## The code chunk
```{r}
#| eval: FALSE
qq <- ggplot(exam_data,
       aes(sample=ENGLISH)) +
  stat_qq() +
  stat_qq_line()
  
sw_t <- exam_data %>%
   shapiro_test(ENGLISH) %>%
   gt()
 
tmp <- tempfile(fileext = '.png')
gtsave(sw_t, tmp)
table_png <- png::readPNG(tmp, native= TRUE)
 
qq + table_png
 

```

:::



